---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# arrayrefs

<!-- badges: start -->
<!-- badges: end -->

The goal of arrayrefs is to get byte range refs with VirtualiZarr but (mostly) without xarray. 

There's very simple information in a reference database, it's path, offset, size, varname, and the (chunk) index. If we get this in an 
simple format we can easily express in different ways for other formats. 



## Installation

Not yet


```{r include = FALSE, eval=FALSE}
#You can install the development version of arrayrefs like so:

remotes::install_github("mdsumner/arrayrefs")
```

## Example

This is a basic example (hey this needs very recent virtualizarr, working on my HPC for now)

```{r example,eval=FALSE,include =TRUE}
library(arrayrefs)
## compare approach here to https://github.com/zarr-developers/VirtualiZarr/issues/533
library(sooty)
ds <- dataset()
ds@id <- "oisst-avhrr-v02r01"
s3files <- gsub("/vsis3/", "s3://", ds@source$source)
## we could just sub this in in place of s3:// but haven't check virtualizarr will allow
s3_creds <-  reticulate::dict(endpoint_url = "https://projects.pawsey.org.au", anon = TRUE)
ro <- reticulate::dict(storage_options = s3_creds)


options(parallelly.fork.enable = TRUE, future.rng.onMisuse = "ignore")
library(furrr); plan(multicore)
vars <- c("sst", "anom", "err", "ice")
dslist <- future_map_dfr(s3files, arrayrefs:::getrefs, vars, reader_options = ro)

arrow::write_parquet(dslist, "examples/oisst.nc.parquet")
```

That took a couple of minutes, no dask problems and no need for concurrent.futures (this is very light on chunks though, only one chunk per file). 
Obviously we wasted storage space with the url and the varname and the index, but we don't have to navigate complex object trees this way. It's a balance. 

```
dslist
# A tibble: 63,688 × 5
   path                                              offset length index varname
   <chr>                                              <int>  <int> <chr> <chr>
 1 s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/dat… 1.03e6 683590 0.0.… sst
 2 s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/dat… 4.77e4 684063 0.0.… anom
 3 s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/dat… 7.32e5 222214 0.0.… err
 4 s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/dat… 9.54e5  70081 0.0.… ice
 5 s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/dat… 1.03e6 684863 0.0.… sst
 6 s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/dat… 4.77e4 686733 0.0.… anom
 7 s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/dat… 7.34e5 223997 0.0.… err
 8 s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/dat… 9.58e5  71560 0.0.… ice
 9 s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/dat… 1.03e6 683768 0.0.… sst
10 s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/dat… 4.77e4 685793 0.0.… anom
# ℹ 63,678 more rows
# ℹ Use `print(n = ...)` to see more rows
> dslist$path[1]
[1] "s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/data/sea-surface-temperature-optimum-interpolation/v2.1/access/avhrr/198109/oisst-avhrr-v02r01.19810901.nc"
> dslist$path[nrow(dslist)]
[1] "s3://idea-10.7289-v5sq8xb5/www.ncei.noaa.gov/data/sea-surface-temperature-optimum-interpolation/v2.1/access/avhrr/202504/oisst-avhrr-v02r01.20250404_preliminary.nc"
length(unique(dslist$path))
[1] 15922

```

## Code of Conduct
  
Please note that the arrayrefs project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
